### 2016-12-23 openadcess.tex

\documentclass{article}

\usepackage{color}
\usepackage{xspace}

% \todo macro
\newcommand{\todo}[1]
{{\bf\color{red}[#1]}}

\newcommand{\premium}{Gold Plus\xspace}

\begin{document}
\section{Introduction}

\section{Related Work}

\input{digital-library}

\section{Paper Sponsorships}

\section{Future Work}
\todo{Crazy dystopian future stuff here}

\end{document}