### 2017-01-29 openadcess.tex

\documentclass[sigtbd]{sigtbd17-style}

\usepackage{amsfonts}
\usepackage{censor}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xspace}

% \todo macro
\newcommand{\todo}[1]
{{\bf\color{red}[#1]}}

\newcommand{\premium}{Gold Plus\xspace}

\begin{document}

\title{\todo{Something} presents OpenAdcess}
\maketitle

\begin{abstract}
  \todo{Abstract}
\end{abstract}

\section{Introduction}

\section{Related Work}

\input{digital-library}

\input{paper-sponsorships}

\input{future-work}

\end{document}