### 2017-02-17 aesthetics.tex
oops, now i really added the aesthetics section

We should have a page that uses pretentious Clement Greenberg-esque language to talk about "perspective" and "foreground" and "background" and "depth of field" and so forth. Most papers are pretty boring: a bird's eye view. But as Freeman Dyson notes, there are birds and frogs in research. And frogs like to get in the mud. And frogs like to croak about Budweiser.

In the foreground of this page-long section, there is a giant can of Budweiser.