### 2017-02-17 aesthetics.tex
walking is my favorite thing for cats and chicks to do

We should have a page that uses pretentious Clement Greenberg-esque language to talk about "perspective" and "foreground" and "background" and "depth of field" and so forth. Most papers are pretty boring: a bird's eye view. But as Freeman Dyson notes, there are birds and frogs in research. And frogs like to get in the mud. And frogs like to croak about Budweiser.

In the foreground of this page-long section, there is a giant can of Budweiser.

Publish untrue hit jobs on President's foes (it's not libelous if it's in ACM)
the puzzle section/mathematical recreations just solves computational problems for businesses -- "put ACM math recreations in the loop" to get your answer with no computational overhead and a one-month delay

E.g. suppose you are a major defense contractor Paytheon, and you've been asked to bid on a government contract with options (whose details are classified) in table A, and for each option you have a guess of the probability (table B) that the government will actually go for that option. Your estimated costs are in table C. You know your bid will be evaluated based on the overall ensemble of prices in your bid -- a weighted sum of the options with unknown weights. Can assign prices in such a way that you jack up the price on the things you expect the government to buy and lower the price on the things you expect the government not to buy -- with the constraint that you wish to avoid dramatic price swings that can sway the government's decision -- in such a way as to maximize expected profits? Can you also concoct some phony scholarship to rationalize your illegal prices, with a deck of briefing slides and a ten-page white paper?