### 2017-02-28 openadcess.tex
Abstract

\documentclass[sigtbd]{sigtbd17-style}

\usepackage{amsfonts}
\usepackage{censor}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xspace}

% \todo macro
\newcommand{\todo}[1]
{{\bf\color{red}[#1]}}

\newcommand{\premium}{Gold Plus\xspace}

\begin{document}

\title{\todo{Something} presents OpenAdcess}
\maketitle

\begin{abstract}
  In this paper we examine an alternate funding model for open access papers.
  Through modest modifications to the ACM Digital Library and the addition of
  paper sponsorships we can make the entire corpus of ACM works free to the
  public.
  We also peer into a chilling future that only we can prevent.
  Will we stop the ACM from encroaching on our everyday freedoms?
  Continue existing and participating in reality to find out!
\end{abstract}

\section{Introduction}

\section{Related Work}

\input{digital-library}

\input{paper-sponsorships}

\input{future-work}

\end{document}