### 2017-02-28 introduction.tex
Took a stab at the intro

\section{Introduction}
Open access is a movement to make papers and other research contributions free
accessible to the general public \todo{cite}.
The ACM has been working for some time to achieve the open access dream but
have found the dream to be prohibitively expensive \todo{cite}.
To fund this venture, we have chosen to adopt the techniques employed by
various other websites and publications that manage to distribute content for
free.

Section~\ref{sec:related} details techniques used by other services to achieve free
downloads and paper distribution.
Section~\ref{sec:lib} describes the upcoming changes to the Digital Library to
bring it up to current standards for download portals.
Section~\ref{sec:sponsors} adapts techniques from tabloids to ensure that papers
themselves are profitable.
Section~\ref{sec:future} describes a future that directly results from the
publication of this paper.
We provide this look into the future so that we may attempt to contain the
beast that is the ACM while still providing free papers to all.