### 2017-03-01 introduction.tex
Hecka citations

\section{Introduction}
Open access is a movement to make papers and other research contributions free
accessible to the general public \cite{oa}.
The ACM has been working for some time to achieve the open access dream but
have found this dream to be prohibitively expensive.
As such, we have required authors to opt-in to open access by paying the cost
themselves \todo{cite (and archive current authors.acm.org page)}.
However, we have found a way to provide open access papers at no cost to the
authors by adopting the techniques employed by various other websites and
publications that manage to distribute content for free.

Section~\ref{sec:related} details techniques used by other services to achieve free
downloads and paper distribution.
Section~\ref{sec:lib} describes the upcoming changes to the Digital Library to
bring it up to current standards for download portals.
Section~\ref{sec:sponsors} adapts techniques from tabloids to ensure that papers
themselves are profitable.
Section~\ref{sec:future} describes a future that directly results from the
publication of this paper.
We provide this look into the future so that we may attempt to contain the
beast that is the ACM while still providing free papers to all.