### 2017-03-01 conclusion.tex
Draft pass

\section{Conclusion}
By bringing ACM up to date with the latest advertising techniques described in
this paper we believe that we can finally provide our papers to the masses for
free and at no cost to the authors.
Unfortunately, our trip to the future indicates that this idea works all too
well.
We release this knowledge onto the world knowing that the next generation must
pay for our free papers with their work in the ACM dungeons and surrender of
their corporations.
We hope that future generations will somehow stop the beast that is the ACM
before it grows too powerful, but our own desire for fame outweighs our desire
to save our children from this madness.