### 2017-03-01 related-work.tex

\section{Related Work}
\label{sec:related}

Download.com \cite{download-com} provides free downloads of popular programs to
the public.
We borrow their idea of using misleading advertisements to trick users into
downloading programs they may not want in \autoref{sec:fake}.

Rapidshare \cite{rapidshare} was a website that provided rate limited free
downloads while simultaneously selling a paid download service that was
superior to the free tier.
This technique was meant to entice users to pay for the service.
Unfortunately, as Rapidshare no longer exists it seems that this technique
alone is insufficient to cover the costs of running a download service.
However, we still employ this technique of upselling the reader by severely
limiting the free tier as a supplemental revenue stream.

Wisdom \cite{wisdom} and many other free rags employ heavy use of in paper ads,
native ads, and classifieds.
We also adopt these techniques to embed in all of our papers in
\autoref{sec:sponsors}.
Additionally, Wisdom claims most holism but we have the narrow computational
tools to prove that we are more holistic.

Edsger W. Dijkstra has previously proposed the revolutionary idea of
capitalizing mathematics \cite{cap-math}.
We expand on this idea with our branded mathematical objects in
\autoref{sec:brands}.