### 2017-04-10 openadcess.tex

\documentclass[sigtbd]{sigtbd17-style}

\usepackage{amsfonts}
\usepackage{censor}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xspace}

% \todo macro
\newcommand{\todo}[1]
{{\bf\color{red}[#1]}}

\newcommand{\premium}{Gold Plus\xspace}

\begin{document}

\title{BurgerTown Presents: Open Adcess}
\maketitle

\begin{abstract}
  In this paper we examine an alternate funding model for open access papers.
  Through modest modifications to the ACM Digital Library and the addition of
  paper sponsorships, we can make the entire corpus of ACM works free to the
  public. What does the future have in store for the open-adcess movement?
  Continue existing and participating in reality to find out!
\end{abstract}

\input{introduction}

\input{related-work}

\input{digital-library}

\input{paper-sponsorships}

\input{future-work}

\input{conclusion}

\bibliography{references}{}
\bibliographystyle{plain}

\end{document}