### 2017-04-23 ideas.tex
current

* one of the links referenced is a terrifying URL that ends in .exe

* Featurette: Does Commercialism Chill Scientific Inquiry (sponsored by Chill's restaurant): Take a look at science -- people valorize metrics, hack p-values, hype results for media coverage and access to scarce research funds, try to boot the ego. Paul Erd\"os may have been right about math but he's wrong about science. Science is a machine for converting sacred principles (openness) and good-to-the-last-drop Maxwell House coffee ("We gave you a Haggadah. It's the least you could do.") into profane, Taylorizable metrics (openness scores). Science is a rat race already. The introduction of incentives for commercial relevance need not constrain the liberty of researchers. Is the pursuit of science that is flattering to the scientist any different than the pursuit of science flattering to the all-new 2017 Jep Liberty? Discuss the latest research about the miracle superfood bacon-wrapped rib nachos at your neighborhood Chill's today!

* Don't think that the easy-to-grasp numbers are necessarily more widely grasped. The ACM children's program, \emph{Shareware Street}, in a disgusting commercial display, infamously allows the numbers with the deepest pockets to sponsor its programming. Consequently, some brands may choose to purchase numbers more abstruse (i.e., admitting no obvious simple English formulation) that they may have money the more with which to bribe \emph{Shareware Street} producers.

* Keep in mind, while sponsors purchase naming rights based on a \emph{description} of a number or set, they are in fact purchasing the number itself. If a speculator purchases, based on a description, a number that turns out not to exist, the speculator is left with nothing, whereas the purchaser of the paradox---the giving a name to a non-existent thing---involving the description, would remain a sponsor of the paradox if it is indeed a valid paradox. If someone purchases, based on description, a number that is already sponsored by another entity, she owes that entity 50 push-ups.