### 2017-04-23 introduction.tex

\section{Introduction}
Open access is a movement to make papers and other research contributions
freely accessible to the general public \cite{oa}.
The ACM has been working for some time to achieve the open access dream but
has found this dream to be prohibitively expensive.
As such, we have required authors to opt-in to open access by paying the cost
themselves \cite{auth}.
However, we have found a way to provide open access papers at no cost to the
authors by adopting the techniques employed by various websites and
publications that manage to distribute content for free.

Section~\ref{sec:related} details techniques used by other services to achieve
free downloads and paper distribution.
Section~\ref{sec:lib} describes the upcoming changes to the Digital Library to
bring it up to the current standards for download portals.
Section~\ref{sec:sponsors} adapts techniques from tabloids to ensure that
the papers themselves are profitable.
Section~\ref{sec:future} provides work from a future that directly results from
a present that includes the present work.